\replaced{Towards an o}{O}pen-source landscape for 3D CSEM modelling

### 2021-01-30 manuscript/Manuscript.tex
Revision Draft; now to co-authors.

A layered model fails to show the strength of 3D modellers, and in reality one would not choose a 3D code to compute responses for a layered model. However, it is one of the few examples that can be verified with semi-analytical results, and is therefore still a valuable test. The results are shown in Figure~\ref{fig:results-layered}, \replaced{in the left column the amplitudes and in the right column the phases}{where the left, middle, and right columns are for the real part, imaginary part, and absolute value, respectively (dashed lines indicate negative values)}. The top row shows the actual, semi-analytically computed responses, for which we used \empymod\citep{GEO.17.Werthmuller}. The second and the third row show the relative percentage error of the inline and the broadside receivers, respectively; note that for the 1D case the two broadside lines for $y=\pm3\,$km are identical.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-layered.png}
  \caption{\itodo{Adjust Figures 1 (Amplitude/Phase) and corresponding text in manuscript.} Results of the layered model: The semi-analytical inline and broadside responses are shown in the top row. The relative percentage error of the four 3D modellers are shown in the second and the third row for the inline and broadside responses, respectively.}
  \label{fig:results-layered}
\end{figure*}
%

\added{The misfit is generally in the order of 1\,\% or less, the biggest misfits are close to the source and where there are rapid phase changes. The errors become substantial towards zero offset, particularly for the results of the FV codes. These comparably significant errors could be eliminated by choosing a finer discretization around the source, at the cost of increased computation time. Another possibility is to have a finer mesh for short offsets, and a coarser mesh for large offsets.}

@@

Note that \emg3d and \simpeg used the same rectilinear mesh in this example. The use of an octree mesh in \simpeg could greatly reduce the size of the mesh and resultant computation time, but at the cost of accuracy, which is reduced when averaging over layer interfaces to coarsen cells and because of the factor-of-two expansion rate of coarsened cells. As the aim of this study is to focus on accuracy of the simulations and compare discretization approaches, we chose to use the same mesh as in the \emg3d simulation. \added{Although we only show the x-directed field in this comparison, the y-directed field yields a similar result with respect to the misfit. A different story in this scenario are the z-directed fields, which is not continues across the boundary on the seafloor. Having receivers close to such a non-continues boundary requires advanced interpolation routine} \citep[e.g., ][]{GJI.15.Shantsev, GP.11.Wirianto}\added{. This is no problem for the used FE codes within their basis functions, but the used FV codes do currently not have such routines implemented. }

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 48 &   232 & 244.1 & 5899578 \\ % dof: 11799156 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   123 &   0.4 & 4813216 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

The block model is a derivation of the \emph{Dublin Test Model~1} from the first EM modelling workshop described by \cite{GJI.13.Miensopust}. We use the same layout of the blocks but adjust the dimensions and resistivities to a typical marine CSEM problem, as shown in Figure~\ref{fig:model-block}. Additionally, we add our \emph{Layered Model} as a VTI background. The three resistive blocks have resistivities of $\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines), $\rho=100\,\ohmm$ (thin plate, South-East), and $\rho=500\,\ohmm$ (cube, North-West).
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/model-block.png}
  \caption{Sketch of the block model, consisting of the layered background model with three resistive blocks, embedded in the thick background layer which has VTI with $\lambda=\sqrt{\rho_\textrm{v}/\rho_\textrm{h}}=\sqrt{2}$.}
  \label{fig:model-block}
\end{figure}
%

There are no semi-analytical solutions for such a model to verify our results, and we can only validate them by comparing different results. For this we use the normalised root-mean square difference (NRMSD) between two responses $R_1$ and $R_2$, given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200\,\frac{|R_1 - R_2|}{|R_1| + |R_2|}\ ,
  \label{eq:nrmsd}
\end{equation}
%
to which we refer as simply the \emph{normalised difference} throughout the paper.

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the normalised differences between the absolute responses of the codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block.png}
  \caption{\itodo{Adjust Figures 3 (\emg3d-\simpeg; \custem-\petgem; FV-FE) and corresponding text in manuscript.}Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the normalised differences (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

A few interesting points stand out:
\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the normalised difference is generally below a few percent, increasing towards larger offsets.
  \item The normalised difference between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low normalised difference is because these codes use the same rectilinear mesh for modelling. The increasing normalised difference towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods.
\end{enumerate}

The corresponding required runtime and memory are listed in Table~\ref{tbl:comp-block}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the block model.}
\label{tbl:comp-block}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 48 &   312 & 281.8 & 6014440 \\ % dof: 12,028,880 / 2  (Real+Imag -> Complex)
  \emg3d  &  1 &   119 &   0.4 & 4813216 \\
  \petgem & 24 &   238 & 152.8 & 2455868 \\
  \simpeg &  4 & 10295 & 280.4 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

The full MR3D model consists of $1022 \times 371 \times 1229$ cells, totalling to almost 466 million cells, where each cell has dimensions of 25 by 75 by 5\,m. The model was upscaled for the computation to 515 by 563 by 310 cells, totalling to almost 90 million cells, where each cell has dimensions of 100 by 100 by 20\,m. Both the full model and the upscaled computational model are released \replaced{as open source}{openly}. The published data set consists of a regular grid of receivers of 20 in eastern direction by 25 in northern direction, 500 receivers in total, with 1\,km spacing located on the irregular seafloor. 45 source-towlines were located on the same grid above each receiver line, 50\,m above the seafloor, with shots every 100\,m. The computed responses for one of the receivers, \texttt{04Rx251a}, are shown in the original publication for the East-West inline-source \texttt{04Tx013a} and the East-West broadside-source \texttt{04Tx014a} (broadside offset of 1\,km). We reproduce the responses for this receiver and corresponding source-lines in our comparison. The $x$-$z$ cross-section of the horizontal resistivity model at the receiver position is shown in Figure~\ref{fig:model-marlim}, together with the receiver and sources positions. All layer have a VTI with $\lambda=\sqrt{2}$, except for the air layer, the seawater, and the salt layer, which are all isotropic.

@@

It is important to note that the published results nor any of our results is the \emph{truth}, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the normalised differences between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2published.png}
  \caption{\itodo{Improve figure.}Normalised differences for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
  \label{fig:results-marlim_2published}
\end{figure*}
%
A few conclusions can be drawn from this figure: In general the normalised difference is roughly 10\,\% or less. The structured meshes have in some cases a smaller normalised difference, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some normalised differences have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}. The origin of this zigzag pattern is most likely the bathymetry.\itodo{Rune confirmed this; adjust} We suspect that the code of the published data did probably use the bathymetry information in addition to the resistivity cube. This would explain why the codes using the tetrahedra mesh, which uses the bathymetry as well, does not show this feature.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the normalised differences between our codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2ours.png}
  \caption{\itodo{Improve figure.}Normalised differences, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
  \label{fig:results-marlim_2ours}
\end{figure*}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their normalised difference is generally below 0.03\,\%. The reason is simple: \custem and \petgem use the exactly same mesh for the computation of this example, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the normalised differences between our codes look similar as in the comparison with the published results, and the principal source of the difference must be in the different discretizations.

@@

The required runtime and memory to compute the shown MR3D responses for the four codes are listed in Table~\ref{tbl:comp-marlim}. The memory requirement varies from \itodo{Check at end again} $0.5-230\,$GiB, and the runtime from roughly 7\,min to 20\,min, where the codes were run on very different machines from servers to supercomputers (see Table \ref{tbl:machines}) and use between 1 and 96 processes in parallel.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the MR3D model.}
\label{tbl:comp-marlim}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 64 &  872 & 230.1 & 1918106 \\ % dof: 3836212 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 & 1246 &   0.5 & 5998992 \\
  \petgem & 96 &  524 & 175.4 & 1918106 \\
  \simpeg &  4 &  422 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

\itodo{SOMEWHERE: Address NRMSD of MR3D results better; EMGS code elan uses bathymetry file; difference not error; very strong NRMSD; although visually identical.}

@@

%% 1. Results: good agreement %%
\itodo{Adjust/check text for amplitude/phase plots.}
Overall, we observed an excellent match between the solutions. The amplitudes of the layered model have mostly a relative error of less than 1--2\,\%. Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses. The cross-validation of the results for the block model yields a similar picture, with normalised differences within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the misfits are higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the normalised difference is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers \replaced{increase}{rise} to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that 5--10\,\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempts were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in \added{the} x- and y-directions for CSEM computations. In order to represent the fine-scale model on a mesh suitable for computation, each of our codes did its own upscaling and extrapolation. However, we were unsuccessful in our attempt to reproduce the published results.  The reason is that there are many ways how you can extrapolate a model, and without knowing the details of how MR3D represents the model on the computational mesh, we could not re-produce this step. We reached out to the authors of MR3D, and they kindly agreed to also publish\deleted{ed} the upscaled and extended computational model. It is only with this model that we were able to obtain comparable results.

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there now exist several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be an ideal scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, the FE codes were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling \citep{GJI.11.Schwarzbach,GJI.16.Key}. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.
\itodo{Add comment: FV mesh creation much easier / faster; FE have then the advantage of better interpolation and easier possibility for adaptive grids.}

>>>>

A layered model fails to show the strength of 3D modellers, and in reality one would not choose a 3D code to compute responses for a layered model. However, it is one of the few examples that can be verified with semi-analytical results, and is therefore still a valuable test. The results are shown in Figure~\ref{fig:results-layered}, \replaced{in the left column the amplitudes and in the right column the phases}{where the left, middle, and right columns are for the real part, imaginary part, and absolute value, respectively (dashed lines indicate negative values)}. The top row shows the actual, semi-analytically computed responses, for which we used \empymod\citep{GEO.17.Werthmuller}. The second and the third row show the relative percentage error of the inline and the broadside receivers, respectively; note that for the 1D case the \added{x-directed electric fields for the} two broadside lines for $y=\pm3\,$km are identical.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-layered.png}
  \caption{Results of the layered model: The semi-analytical inline and broadside responses are shown in the top row. The relative percentage error of the four 3D modellers are shown in the second and the third row for the inline and broadside responses, respectively.}
  \label{fig:results-layered}
\end{figure*}
%
\added{The misfit is generally in the order of 1\,\% or less, the biggest misfits are close to the source and where there are rapid phase changes.}

@@

\itodo{Expand and adjust, wait for final result \simpeg (same mesh or different mesh?); some of this might be outdated.}
Note that \emg3d and \simpeg used the same rectilinear mesh in this example. The use of an octree mesh in \simpeg could greatly reduce the size of the mesh and resultant computation time, but at the cost of accuracy, which is reduced when averaging over layer interfaces to coarsen cells and because of the factor-of-two expansion rate of coarsened cells. As the aim of this study is to focus on accuracy of the simulations and compare discretization approaches, we chose to use the same mesh as in the \emg3d simulation. \added{Although we only show the x-directed field in this comparison, the y-directed field yields a similar result with respect to the misfit. A different story in this scenario is the z-directed field, which is not continues across the boundary on the seafloor. Having receivers close to such a non-continues boundary requires advanced interpolation routine} \citep[e.g., ][]{GJI.15.Shantsev, GP.11.Wirianto}\added{. This is no problem for the used FE codes within their basis functions, but the used FV codes do currently not have such routines implemented. The grid for the FV codes in this first example with a focus on precision is created in such a way that  the receivers are located on the corresponding nodes, so no interpolation is required to obtain the responses.}

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 48 &   232 & 244.1 & 5899578 \\ % dof: 11799156 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   216 &   0.5 & 6004144 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

The block model is a \replaced{CSEM adaption}{derivation} of the \added{MT} \emph{Dublin Test Model~1} from the first EM modelling workshop described by \cite{GJI.13.Miensopust}. We use the same layout of the blocks but adjust the dimensions and resistivities to a typical marine CSEM problem, as shown in Figure~\ref{fig:model-block}. Additionally, we add our \emph{Layered Model} as a VTI background. The three resistive blocks have resistivities of $\rho=10\,\ohmm$ (shallow beam perpendicular to survey lines), $\rho=100\,\ohmm$ (thin plate, South-East), and $\rho=500\,\ohmm$ (cube, North-West).
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/model-block.png}
  \caption{Sketch of the block model, consisting of the layered background model with three resistive blocks, embedded in the thick background layer which has VTI with $\lambda=\sqrt{\rho_\textrm{v}/\rho_\textrm{h}}=\sqrt{2}$.}
  \label{fig:model-block}
\end{figure}
%

There are no semi-analytical solutions for such a model to verify our results, and we can only validate them by comparing different results. For this we use the normalised root-mean square difference (NRMSD) between two responses $R_1$ and $R_2$, given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = \replaced{100\,\frac{|R_1 - R_2|}{(|R_1| + |R_2|)/2}}
                             {200\,\frac{|R_1 - R_2|}{|R_1| + |R_2|}}\ ,
  \label{eq:nrmsd}
\end{equation}
%
to which we refer as simply the \emph{normalised difference} throughout the paper.

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the normalised differences between the absolute responses of the codes. \added{Note that for visual reasons we only show the normalised differences between the two FV codes, the two FE codes, and between \emg3d and \custem as an example of a cross-FV-FE comparison; the three other cross-method comparisons look quantitatively similar.}
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block.png}
  \caption{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the normalised differences (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

A few interesting points stand out:
\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the normalised difference is generally below a few percent, increasing towards larger offsets.
  \item \itodo{Adjust with final \simpeg results; conclusion does currently not hold.} The normalised difference between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low normalised difference is because these codes use the same rectilinear mesh for modelling. The increasing normalised difference towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods.
\end{enumerate}

The corresponding required runtime and memory are listed in Table~\ref{tbl:comp-block}.
\ymark{We could remove Table 4, as \#procs and \#doc are equal, and runtime and memory are almost identical for all codes in these two examples. What do you think? I would then replace it in the following way:
«The corresponding required runtime and memory are \added{almost identical to the ones for the layered model} listed in Table~\replaced{\ref{tbl:comp-layered}}{\ref{tbl:comp-block}}.»}
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the block model.}
\label{tbl:comp-block}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 48 &   312 & 281.8 & 6014440 \\ % dof: 12,028,880 / 2  (Real+Imag -> Complex)
  \emg3d  &  1 &   213 &   0.5 & 6004144 \\
  \petgem & 24 &   238 & 152.8 & 2455868 \\
  \simpeg &  4 & 10295 & 280.4 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

The full MR3D model consists of $1022 \times 371 \times 1229$ cells, totalling to almost 466 million cells, where each cell has dimensions of $25$\replaced{$\times$}{by}$75$\replaced{$\times$}{by}$5$\,m. The model was upscaled for the computation to $515$\replaced{$\times$}{by}$563$\replaced{$\times$}{by}$310$ cells, totalling to almost 90 million cells, where each cell has dimensions of $100$\replaced{$\times$}{by}$100$\replaced{$\times$}{by}$20$\,m. Both the full model and the upscaled computational model are released \replaced{as open source}{openly}. The published data set consists of a regular grid of receivers of 20 in eastern direction by 25 in northern direction, 500 receivers in total, with 1\,km spacing located on the irregular seafloor. 45 source-towlines were located on the same grid above each receiver line, 50\,m above the seafloor, with shots every 100\,m. The computed responses for one of the receivers, \texttt{04Rx251a}, are shown in the original publication for the East-West inline-source \texttt{04Tx013a} and the East-West broadside-source \texttt{04Tx014a} (broadside offset of 1\,km). We reproduce the responses for this receiver and corresponding source-lines in our comparison. The $x$-$z$ cross-section of the horizontal resistivity model at the receiver position is shown in Figure~\ref{fig:model-marlim}, together with the receiver and sources positions. All layer have a VTI with $\lambda=\sqrt{2}$, except for the air layer, the seawater, and the salt layer, which are all isotropic.

@@

It is important to note that the published results nor any of our results is the \emph{truth}, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the normalised differences between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2published.png}
  \caption{Normalised differences for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
  \label{fig:results-marlim_2published}
\end{figure*}
%
A few conclusions can be drawn from this figure: In general the normalised difference is roughly 10\,\% or less. The structured meshes have in some cases a smaller normalised difference, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some normalised differences have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}. The origin of this zigzag pattern is most likely the bathymetry. We suspect that the code of the published data did probably use the bathymetry information in addition to the resistivity cube. This would explain why the codes using the tetrahedra mesh, which uses the bathymetry as well, does not show this feature.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the normalised differences between our codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2ours.png}
  \caption{Normalised differences, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
  \label{fig:results-marlim_2ours}
\end{figure*}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their normalised difference is generally below 0.03\,\%. \added{This is why they are shown as one case in Figure~\ref{fig:results-marlim_2published}.} The reason is simple: \custem and \petgem use the exactly same mesh for the computation of this example, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the normalised differences between our codes look similar as in the comparison with the published results, and the principal source of the difference must be in the different discretizations.

@@

The required runtime and memory to compute the shown MR3D responses for the four codes are listed in Table~\ref{tbl:comp-marlim}. The memory requirement varies from $0.5-230\,$GiB, and the runtime from roughly 7\,min to 20\,min, where the codes were run on very different machines from servers to supercomputers (see Table \ref{tbl:machines}) and use between 1 and 96 processes in parallel.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the MR3D model.}
\label{tbl:comp-marlim}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 64 &  872 & 230.1 & 1918106 \\ % dof: 3836212 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 & 1246 &   0.5 & 5998992 \\
  \petgem & 96 &  524 & 175.4 & 1918106 \\
  \simpeg &  4 &  422 &  12.8 &  720146 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

@@

%% 1. Results: good agreement %%
Overall, we observed an excellent match between the solutions. The amplitudes \added{and phases} of the layered model have mostly a relative error of less than 1--2\,\%. \deleted{Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses.} The cross-validation of the results for the block model yields a similar picture, with normalised differences within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the misfits are higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the normalised difference is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers \replaced{increase}{rise} to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that 5--10\,\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempts were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in \added{the} x- and y-directions for CSEM computations. In order to represent the fine-scale model on a mesh suitable for computation, each of our codes did its own upscaling and extrapolation. However, we were unsuccessful in our attempt to reproduce the published results.  The reason is that there are many ways how you can extrapolate a model, and without knowing the details of how MR3D represents the model on the computational mesh, we could not re-produce this step. We reached out to the authors of MR3D, and they kindly agreed to also publish\deleted{ed} the upscaled and extended computational model. It is only with this model that we were able to obtain comparable results.

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there now exist several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be an ideal scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, the FE codes were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling \added{\mbox{\citep{GJI.11.Schwarzbach,GJI.16.Key}}}. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.
\ymark{I added this last sentence. Let me know what you think, we can take it out again or adjust/modify it:}
\added{Also, creating FV meshes is much simpler than creating FE meshes. However, having a FE mesh yields a lot o flexibility with regards to interpolation and refinement and adaption, making it more powerful once created.}
